Dynamical modeling of tidal streams

### 2013-12-31 tex/ms.tex
streams: do streams follow orbit

>>>>

The answer to this question is clearly no. The real question therefore
is how bad the approximation that streams follow orbits really is. The
first question is what is meant by the approximation of streams as a
single orbit. \citet{Sanders13a} define this as meaning that the
frequency differences along the orbit lie in the direction of the
progenitor frequency $\veco_p$. However, this does not faithfully
represent how orbit fitting works in practice
\citep[\eg,][]{Koposov10a}. Orbit fitting works by fitting a
\emph{single} orbit to a stream, such that there are no frequency
differences along the orbit which can be compared with
$\veco_p$. Orbit fitting is more correctly represented by stating that
$t_d$ is infinite in the framework of this paper and using the stream
track to fit data. In this case, the mean $\Delta \veco$ is constant
along the stream and the stream track calculated as in
\sectionname~\ref{sec:track} will be a single orbit.

In our stream model the distribution of $\Delta \veco$ is constant
with time, such that stream members on average are removed to the same
orbit. However, the finite disruption time means that for stream
members to have reached the largest angle offsets from the progenitor,
they have to have been removed onto an orbit that is separated from
the progenitor's by a larger $\Delta \veco$ than average. In our model
this is explicitly expressed by the distribution $p(\Delta
\opar|\Delta \apar)$, calculated in \equationname~(\ref{eq:poparapar})
and shown in the top left panel of
\figurename~\ref{fig:gd1_apar}. Thus, the average orbit changes as a
function of positon along the stream, whether expressed as $\Delta
\apar$, $X$, or $l$, and a single-orbit approximation is necessarily
wrong over the full length of the stream.

It is clear from \figurename~\ref{fig:gd1_apar} that a large part of
the stream near the progenitor, approximately $\Delta \apar < \Delta
\Omega^m\,t_d$, is on average on the same orbit, such that it could be
fit with a single orbit. However, for constraining the gravitational
potential with a tidal stream, it is essential to use as large an arc
of the stream as possible, such that limiting the stream to a small
region around the progenitor to be fit with a single orbit is not
optimal.

Because the orbit-fitting definition used by \citet{Sanders13a} does
not simulate real orbit fitting, the conclusions of their analysis as
regards biases in the recovered potential from the faulty single-orbit
assumption should be treated with caution. While orbit fitting is
wrong, how biased inferences based on it are is therefore still an
open question. However, orbit-fitting-based constraints on the Milky
Way halo's potential based on streams such as GD-1 or the Orphan
stream must be treated with suspicion. The fact that the progenitors
of these streams have not been located means that the observed stream
segments are likely far from the progenitor, where the average orbit
changes as a function of angle along the stream (\ie, right of the
dashed line in \figurename~\ref{fig:gd1_apar}).
